---
title: ccleR6
author: "Vincent J. Carey, stvjc at channing.harvard.edu based on work of Phil Chapman at CRUK"
date: "July 2015"
output:
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{xyz}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```
```{r getp,echo=FALSE,results="hide"}
Sys.setenv("CCLEDB_PATH"="/Users/stvjc/Research/CCLE_CHAPMAN/CellLineData.db")
suppressPackageStartupMessages({
library(ccleR6)
})
```

# Introduction

A large indexed SQLite database has been created by Phil Chapman
to represent CCLE, Achilles, and other integrative data sources
relevant to cancer biology.  

This document describes some
approaches to user interface design.  We indicate how to

- query the database directly
- provide R-level support for substantively useful self-description
- carry out some exemplary 'analyses'

For this code to work you need to have the environment variable
`CCLEDB_PATH` defined to give the path to the SQLite file.

# An R6 interface

I believe that an object that is somewhat fleshed out relative to
the database view provided by dplyr will come in handy.  Therefore
I defined a reference class and have lightly populated it with
some identifier vectors.

```{r getit}
ccle = ccledb$new(.ccleSrc)
ccle
"BRAF" %in% ccle$cngenes  # copy number gene list
```

I have not done much with this yet but I think this will
help with subsetting and shiny query support.

# dplyr-based interaction

We can get a feel for the contents with some simple
commands.  `.ccleSrc` is a globally defined dplyr `src`.
```{r dodp}
library(ccleR6)
.ccleSrc
.ccleSrc %>% tbl("ccle_cell_line_info") %>% head()
```

# Gene annotation

Expression data are provided at probe-set level, with symbols.

```{r lkex}
.ccleSrc %>% tbl("ccle_exprs_tall") %>% head()
```

# Compound annotation

```{r lkdr}
.ccleSrc %>% tbl("ccle_drug_data") %>% head()
```
