---
title: ccleR6 get functions
author: "Vincent J. Carey, stvjc at channing.harvard.edu based on work of Phil Chapman at CRUK"
date: "July 2015"
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{xyz}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```
```{r getp,echo=FALSE,results="hide"}
Sys.setenv("CCLEDB_PATH"="~/BigData/CellLineData/CellLineData.db")
suppressPackageStartupMessages({
library(ccleR6)
})
```

# Introduction

A large indexed SQLite database has been created by Phil Chapman
to represent CCLE, Achilles, and other integrative data sources
relevant to cancer biology.  

This document describes some
approaches to user interface design.  We indicate how to

- query the database directly
- provide R-level support for substantively useful self-description
- carry out some exemplary 'analyses'

For this code to work you need to have the environment variable
`CCLEDB_PATH` defined to give the path to the SQLite file. ie 

```Sys.setenv("CCLEDB_PATH"="path/to/db")```

# An R6 interface

I believe that an object that is somewhat fleshed out relative to
the database view provided by dplyr will come in handy.  Therefore
I defined a reference class and have lightly populated it with
some identifier vectors.

```{r getit}
ccle = ccledb$new(.ccleSrc)
ccle
```

This can be serialized, but the database connection needs to be
refreshed on load.

The "guide vectors" are created as hints on vocabularies in use.
```{r chkg}
"BRAF" %in% ccle$cngenes  # copy number gene list
```

# Some guided filters

We have defined a number of filter functions to
simplify common subsetting operations.
```{r lkfs}
ccle$src %>% filter_compound("Irinotecan")
ccle$src %>% filter_organ("breast")
ccle$src %>% filter_Histology_patt("%wing%")
```

# Query speed

microbenchmark?

# Get Data Function Family

I extended the reference class itself to contain a query set that can be applied across different data sets.

```{r queryspec}
ccle$query_symbols <- c('PTEN', 'BRAF', 'NRAS', 'KRAS', 'SMARCA4', 'PIK3CA')
ccle$query_cell_lines <- (ccle$src %>% filter_organ("skin") %>%  as.data.frame(n=-1))$CCLE_name
ccle$query_compounds <- c('PLX4720', 'AZD6244', 'Lapatinib')
```

Data can then be retrieved by running a function against the R6 class

```{r getdata} 
affy_data <- get_affy(ccle)
head(affy_data)
hybcap_data <- get_hybcap(ccle)
head(hybcap_data)
resp_data <- get_response(ccle)
head(resp_data)
```

Importantly, since the data structure is standardised we can now combine disparate data types into a common data frame

```{r getdata2}
combined_data <- bind_rows(affy_data,hybcap_data,resp_data)
```

# Modelling drug response

The ```get_data``` functions can be combined to give an output suitable for modeling using the ```make_df``` function:

```{r make_df} 
my_df <- make_df(ccle)
my_df[1:5,1:10]
summary(lm(AZD6244_resp ~ as.factor(BRAF_hybcap), data=my_df))
summary(lm(AZD6244_resp ~ as.factor(BRAF_hybcap) + as.factor(NRAS_hybcap), data=my_df))
```

# Visualising drug response

Finally a convenient heatmap plot can be generated using the ```make_heatmap``` function to visualise drug response vs genetic features.

```{r make_heatmap}
make_heatmap(ccle, compound='AZD6244')
```

# Widgets

Two displays of the Barettina paper (Figures 4a [lower dotplot
with sensitivity against line] and 4c)
can be approached interactively using the multiWidget function.
Briefly, `multiWidget(ccle)` will generate a browser window
with two panes, concatenated on one page.  The first one allows
selection of a gene for which hybrid capture mutation information
has been obtained, and a compound.  The IC50s for cell lines
with hybrid capture data are ordered and displayed.
The second panel allows selection of
a gene whose expression will be summarized across primary tumor sites.

```{r shdem,fig=TRUE,echo=FALSE,fig.height=4.6}
library(grid)
library(png)
im = readPNG("images/shinyDemo.png")
grid.raster(im)
```

# To do

The vocabularies used for cell lines and tumor anatomy are complicated
and need harmonization and streamlining.

Full dose-response information is available and should be exposed.

Interfaces for filters and joins need to be specified and deployed for
common use cases.

Additional interactivity such as tooltips over points that can 
describe mutation profiles.
